### 2020-05-20 paper/icsme20.tex
new paper version

\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{booktabs}
\usepackage{framed}
\usepackage{float}
\usepackage{xcolor}
\usepackage{footnote}
\usepackage{xfrac}
\usepackage{balance}
\usepackage{hyperref}
\usepackage{array}
\usepackage{subfigure}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{listings}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{lightgreen}{rgb}{0.6,0.9,0.6}
\definecolor{lightyellow}{rgb}{0.9,0.9,0.6}
\definecolor{lightorange}{rgb}{0.9,0.8,0.6}
\definecolor{lightred}{rgb}{0.9,0.7,0.7}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{lightgray}{rgb}{0.8,0.8,0.8}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\usepackage{pifont}
\newcommand{\lstbg}[3][0pt]{{\fboxsep#1\colorbox{#2}{\strut #3}}}

\lstdefinestyle{CStyle} {
    language=C,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

\lstdefinestyle{JavaStyle} {
    language=Java,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
    numbers=left,
    stepnumber=1,
    xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

\lstdefinestyle{PHPStyle} {
    language=PHP,
    alsolanguage=HTML,
    backgroundcolor=\color{white},   % choose the background color
    basicstyle=\ttfamily\scriptsize,        % size of fonts used for the code
    breaklines=true,                 % automatic line breaking only at whitespace
    captionpos=b,                    % sets the caption-position to bottom
    commentstyle=\color{mygray}\bfseries,    % comment style
    escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
    keywordstyle=\color{blue},       % keyword style
    stringstyle=\color{mymauve},     % string literal style
    frame=single,
	numbers=left,
    stepnumber=1,
	xleftmargin=2em,
    escapeinside={/*!}{!*/},
    moredelim=**[l][\color{mygreen}]{+\ },
    moredelim=*[l][\color{red}]{-\ }
}

\newcounter{lstannotation}
\setcounter{lstannotation}{0}
\renewcommand{\thelstannotation}{\ding{\number\numexpr181+\arabic{lstannotation}}}
\newcommand{\annotation}[1]{\refstepcounter{lstannotation}\label{#1}\thelstannotation}

%
% Add comments in the text
%
\newboolean{showcomments}
\setboolean{showcomments}{true}
%\setboolean{showcomments}{false}

\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[3]{
  {\color{#2}\small\fbox{\bfseries\sffamily\scriptsize#1}}
  {\color{#2}\sffamily\small$\triangleright~$\textit{\small #3}$~\triangleleft$}
  }
  }
  {\newcommand{\nb}[3]{}
  }

\newcommand\Sof[1]{\nb{Sofia}{red}{#1}}
\newcommand\Luis[1]{\nb{Luis}{mygreen}{#1}}
\newcommand\Rui[1]{\nb{Rui}{blue}{#1}}

\makeatletter
\newcommand\footnoteref[1]{\protected@xdef\@thefnmark{\ref{#1}}\@footnotemark}
\makeatother

\begin{document}

% \title{Fixing Vulnerabilities Hinders Maintainability}
\title{\Sof{TBA - New Titles in the comments}}
%\title{Coding practices are important to produce secure software: Teach them to your students and developers!}
%\title{The reason why you should be applying coding practices: safer software.}
%\title{What is the impact of security patches in software evolution?}
%\title{Hey, developers! You should pay more attention to your patches.}
%\title{Are good coding pratices used when patching software vulnerabilities?}

\author{
    Anonymou(s) Author(s)

%     \IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address}
% \and
% \IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address}
% \and
% \IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address}
}

\maketitle

\begin{abstract}
Security is a requirement of utmost importance to produce software with 
high quality. However, there is still a considerable amount of 
vulnerabilities being discovered and fixed, almost weekly. We suspect 
that when developers address these vulnerabilities on their codebases, 
they actually affect the maintainability of software, potentially 
introducing other vulnerabilities. This paper evaluates the impact of 
patches to improve security on the maintainability of open-source software. 
Maintainability is measured based on the Better Code Hub’s model of 10 
guidelines using a dataset including 1335 security commits.
Results show that fixing software vulnerabilities may hinder the maintainability 
of open-source software and that different levels of attention should be given to 
different severity levels, programming languages and types of weaknesses. We 
conclude that changes to code bases to patch 
vulnerabilities need to be performed with extra care and that security-related 
code evolution should be subject in computer curricula.
\end{abstract}

\begin{IEEEkeywords}
Security, Software Maintenance, Open-Source Software
\end{IEEEkeywords}

\section{Introduction}
%
Software quality is important because it is ultimately related to the overall
cost of developing, extending and maintaining software applications~\cite{slaughter1998evaluating}. 
Software quality characteristics include, but are not limited to functional correctness,
reliability, usability, maintainability, and security. Security is an 
important non-functional requirement during the development of software systems. 
However, there is still a considerable amount of vulnerabilities being
discovered, and fixed, almost weekly as disclosed by the Zero Day Initiative
website\footnote{Zero Day Initiative website available at
\url{https://www.zerodayinitiative.com/advisories/published/} (Accessed on \today{})}.
These vulnerabilities are gateways for attackers to penetrate systems and 
exploit their resources. As an example, at the beginning of $2020$, academics published a 
new execution attack (CacheOut) capable of leaking data from Intel CPUs across many 
security boundaries~\cite{cacheOut} $-$ a processor bug that the company attempted to 
address in previous generations and that makes every user using a CPU released before Q$4$ $2018$ vulnerable. Although not a software bug, software can mitigate this type of issues
at the cost of features and/or performance.
%

In $2011$, the International Organization for Standardization (ISO) issued an
update for software product quality ISO/IEC 25010 considering
\emph{Security} as one of the main software product quality characteristics~\cite{iso:2011}. 
However, many developers still lack the knowledge about best practices to deliver secure and high-quality software~\cite{Pothamsetty:2005:SEL:1107622.1107635, 8077802}.
Unmaintainable code is hard to test, analyze and re-use, and thus more difficult
to find vulnerabilities and defects. Therefore, writing maintainable code
supports the production of more reliable software preventing the introduction of 
faults/vulnerabilities in the code bases. ISO describes software 
maintainability as ``the degree of effectiveness and
efficiency with which a software product or system can be modified to improve
it, correct it or adapt it to changes in environment, and in requirements'' on
software quality ISO/IEC 25010~\cite{iso:2011}. Novel tools have been built to 
detect automatically software vulnerabilities (e.g., FindBugs, Infer and more). 
Despite these efforts, improving software security is not a trivial task and requires 
implementing patches that might affect software maintainability. 
Our suspiciousness is that some of the these patches may have a negative 
impact on the software maintainability and, possibly, even be the cause of the 
introduction of new vulnerabilities $-$ harming software reliability. Therefore, 
in this paper, we present an empirical study on the impact of patches of 
vulnerabilities on software maintenance across open-source software.
%
% Software maintainability can also be defined as 
% the property of software that reveals how easily the software system can be 
% maintained~\cite{IEEEComputerSociety:2014:GSE:2616205} or recover after a failure. 
% Hedegus et al.
% ($2018$)~\cite{HEGEDUS2018313}, analyzed the difference of maintainability between
% groups of refactored elements and non-refactored and conclude that the source
% code elements subjected to refactorings had significantly lower maintainability
% than the elements not modified.
% There is still little information about the impact of patching vulnerabilities
% in open-source software maintainability. \textit{Are developers affecting software
% maintainability when trying to improve software security?}, it is what we want to
% investigate.
%

As ISO does not provide any specific guidelines/formulas to calculate 
maintainability, we resort to Software Improvement Group (SIG)'s web-based source 
code analysis service Better Code Hub (BCH) to compute the software compliance 
with a set of $10$ guidelines/metrics to produce quality software 
based in ISO/IEC $25010$~\cite{Visser:2016:OREILLY}. The code metrics used by BCH 
were empirically validated in previous work~\cite{Bijlsma:2012:FIR:2317098.2317124, 8530041, cruz2019energyoriented}. 
There are other well-known standards and models that have been proposed
to increase software security: Common Criteria~\cite{common:2009} which received
negative criticism regarding the costs associated and poor technical evaluation;
the OWASP Application Security Verification Criteria~\cite{oswap:2009} which is
focused only on web applications, and a model proposed by Xu et al. ($2013$)
~\cite{6616351} for rating software security (arguably, it was one of the
first steps taken by SIG to introduce security on their maintainability model). 
%
We use BCH's toolset to provide an assessment of maintainability in 
software mainly due to the broad number of metrics used.

From a methodological 
perspective, we leveraged a dataset of $1335$ security patches collected from open-source 
software. We calculate software maintainability before and after the patch to measure its 
impact. This empirical study presents evidence that changes 
applied in the code bases to patch vulnerabilities affect code maintainability.
Results also show that different levels of attention should be given to different
severity levels, programming languages and types of weaknesses.
Little information is known about the impact of security patches on the software
maintainability. With this study, 
we intend to highlight the need for tools to predict the impact of patches on 
maintainability~\cite{4724577}, documentation to help developers adopt security 
best practices~\cite{6311252, 7927935, MESQUIDA201519} and the importance of integrating
security-related code evolution in computer curricula.
%
%BCH compiles the ISO/IEC 25010 into a set of $10$ guidelines that are the
%base of BCH's evaluation. Two examples are the \emph{Write Simple Units of Code}
%based on the McCabe Complexity~\cite{1702388} and \emph{Keep your code base Small}
%based on the idea that smaller code bases are easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}.
%

The main contributions of the present work are:
%
\begin{itemize}
	\item An empirical study on the impact of security changes on software
	maintainability and what patterns need more attention.
	\item A replication package with all scripts and data created to perform the
	empirical evaluation, for reproducibility. Available online:
  \url{https://figshare.com/s/4861207064900dfb3372}.
\end{itemize}
%-
This paper is structured as follows: section~\ref{sec:motivation} introduces an
example of a security refactoring of a known vulnerability found in the
protocol implementation of OpenSSL\footnote{\label{openssl}OpenSSL is a toolkit that
contains open-source implementations of the SSL and TLS cryptographic
protocols. Repository available at \url{https://github.com/openssl/openssl}
(Accessed on \today{})}; section~\ref{sec:methodology} describes the
methodology used to answer the research questions; section~\ref{sec:results}
presents the results and discusses their
implications; section~\ref{sec:threats} enumerates the threats to the validity of
this study; section~\ref{sec:rw} describes the different work and existing
literature in the field of study; and, finally, section~\ref{sec:conclusions}
concludes the main findings and elaborates on future work.
%

\section{Motivation and Research Questions}\label{sec:motivation}
%
Due to time-to-market pressure and the lack of security expertise~\cite{8077802}, code-related
security flaws are generally detected \textit{after the fact}, i.e., when
hackers exploit them. It turns out that fixing these issues is as simple as
modifying the software code base. However, these patches may have a
negative impact on software maintenance --- e.g., developers might follow the
quickest solution to fix the problem and not necessarily the most elegant and
performant one. Thus, we want to understand how security patches impact
the maintainability of software.

As an example, consider the patch of the TLS state machine protocol implementation
in OpenSSL\footnoteref{openssl} to address a memory leak flaw found in the way 
OpenSSL handled TLS status request extension data during session 
renegotiation, and where a malicious client could cause a Denial-of-Service
(DoS) attack via large OCSP Status Request extensions - if OCSP stapling support
was enabled. Online Certificate Status Procotol (OCSP) stapling, formally known as the TLS Certificate 
Status Request extension, is a standard for checking the revocation status of
certificates.

This vulnerability is listed at the Common
Vulnerabilities and Exposures dictionary as CVE-$2016$-$6304$~\footnote{CVE-$2016$-$6304$
details available at \url{http://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2016-6304}
(Accessed on \today{})}. It is amongst the vulnerabilities studied in our
research. The snippet, in Listing~\ref{lst:vuln}, presents the changes performed on the
\emph{ssl/t1$\_$lib.c} file\footnote{CVE-$2016$-$6304$ fix available  at
\url{https://github.com/openssl/openssl/commit/e408c09bbf7c3057bda4b8d20bec1b3a7771c15b}
(Accessed on \today{})} by the OpenSSL developers to patch the vulnerability. 
Every SSL/TLS connection begins with a handshake 
which is responsible for the negotiation between the two parties. 
The OSCP Status Request extension allows the client to verify
the server certificate and enables a TLS server to include
its response in the handshake. The problem in CVE-$2016$-$6304$
is a flaw in the logic of OpenSSL that does not handle memory efficiently
when large OCSP Status Request extensions are sent each time a client 
requests renegotiation. This was possible because the OCSP responses IDs 
were not released between handshakes. Instead, they would be allocated 
again and again. Thus, if a malicious client does it several times it may lead to 
an unbounded memory growth on the server and, eventually, lead to a 
DoS attack through memory exhaustion.

%We remove any OCSP_RESPIDs from a previous handshake
%to prevent unbounded memory growth - CVE-2016-6304

\medskip
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={CStyle}, caption={Patch provided by OpenSSL developers to the
\\CVE-2016-6304 vulnerability on file ssl/t1\_lib.c},label={lst:vuln}]
static int ssl_scan_clienthello_tlsext(SSL *s, PACKET *pkt, int *al){ 
 // [snip]
+  sk_OCSP_RESPID_pop_free(s->tlsext_ocsp_ids, OCSP_RESPID_free); /*!\annotation{lst:func1}!*/
+  if (PACKET_remaining(&responder_id_list) > 0) { 
+     s->tlsext_ocsp_ids = sk_OCSP_RESPID_new_null();
+     if (s->tlsext_ocsp_ids == NULL) { /*!\annotation{lst:func2}!*/
+        *al = SSL_AD_INTERNAL_ERROR;
+        return 0;
+     }
+  } else {
+     s->tlsext_ocsp_ids = NULL;
+  }

   while (PACKET_remaining(&responder_id_list) > 0) {
     OCSP_RESPID *id;
     PACKET responder_id;
     const unsigned char *id_data;
     if (!PACKET_get_length_prefixed_2(&responder_id_list, &responder_id) || PACKET_remaining(&responder_id) == 0) {
          return 0;
     }

-  if (s->tlsext_ocsp_ids == NULL 
-      && (s->tlsext_ocsp_ids = 
-      sk_OCSP_RESPID_new_null()) == NULL) { /*!\annotation{lst:func3}!*/
-    *al = SSL_AD_INTERNAL_ERROR;
-    return 0;
-  }

 // [snip]
 }
\end{lstlisting}

% \medskip
% \setcounter{lstannotation}{0}
% \begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
% \\CVE-2016-6307 vulnerability for file ssl/statem/statem.c},label={lst:vuln}]
%  static SUB_STATE_RETURN read_state_machine(SSL *s){
%  // [snip]
% +   if (!SSL_IS_DTLS(s) && s->s3->tmp.message_size > 0
% +        && !BUF_MEM_grow_clean(s->init_buf,
% +                                (int)s->s3->tmp.message_size
% +                                + SSL3_HM_HEADER_LENGTH)) { /*!\annotation{lst:func2}!*/
% +         ssl3_send_alert(s, SSL3_AL_FATAL, SSL_AD_INTERNAL_ERROR);
% +         SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
% +          return SUB_STATE_ERROR;
%  // [snip]
%  }
% \end{lstlisting}
%
% \medskip
% \setcounter{lstannotation}{0}
% \begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
% \\CVE-2016-6307 vulnerability for file ssl/statem/statem\_lib.c},label={lst:vuln2}]
%  int tls_get_message_header(SSL *s, int *mt){
%  // [snip]
%   l = RECORD_LAYER_get_rrec_length(&s->rlayer) +
%         SSL3_HM_HEADER_LENGTH;
% - if (l && !BUF_MEM_grow_clean(s->init_buf, (int)l)) { /*!\annotation{lst:func3}!*/
% -      SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
% -      goto err;
%    }
%  // [snip]
%
% - if (l && !BUF_MEM_grow_clean(s->init_buf, (int)l +SSL3_HM_HEADER_LENGTH)) {  /*!\annotation{lst:func4}!*/
% -     SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
% -     goto err;
% - }
%
% \end{lstlisting}
%
%\medskip
%\setcounter{lstannotation}{0}
%\begin{lstlisting}[style={CStyle}, caption={Fix provided by OpenSSL developers to the
%\\CVE-2016-6307 vulnerability for file ssl/statem/statem_lib.c },label={lst:vuln}]
% int tls_get_message_header(SSL *s, int *mt){
% // [snip]
%+   if (!SSL_IS_DTLS(s) && s->s3->tmp.message_size > 0
%+        && !BUF_MEM_grow_clean(s->init_buf, 
%+                                (int)s->s3->tmp.message_size
%+                                + SSL3_HM_HEADER_LENGTH)) {
%+         ssl3_send_alert(s, SSL3_AL_FATAL, SSL_AD_INTERNAL_ERROR);
%+         SSLerr(SSL_F_TLS_GET_MESSAGE_HEADER, ERR_R_BUF_LIB);
%+          return SUB_STATE_ERROR;
% // [snip]
% }
%\end{lstlisting}

The code changes performed to patch the vulnerability were:

~\ref{lst:func1} The \texttt{sk\_OCSP\_RESPID\_pop\_free} function removes any
memory allocated to the OCSP response IDs (OCSP$\_$RESPIDs) from a previous handshake to prevent unbounded memory growth - which was not being perfomed before.
After releasing the unbounded memory, the program performs the same check 
in ~\ref{lst:func2} that used to perform in ~\ref{lst:func3} which is responsible 
for handling the application when no OCSP response IDS were being allocated. Now, instead of performing it after packaging processing, it does it previously.

Although performing this patch, seems an elementary problem to repair,
the example shows that there was a significant amount of changes performed
in the code base which performed a negative impact in the software maintanibility.
While patching, the developer introduced $6$ new lines in a method already with 
a large number of lines of code and introduced more complexity to the code
with new branch points which disrupts two of the guidelines proposed by the Software Improvement Group (SIG) for
building maintainable software~\cite{Visser:2016:OREILLY} - \emph{Write Short Units of Code}
and \emph{Write Simple Units of Code}.

% \Sof{The \texttt{read\_state\_machine} function, presented in Listing \ref{lst:vuln}, abstracts a sub-state-machine
% to handle the messages exchanged between peers. When messages are larger than
% what is defined by default for TLS, it may trigger a memory corruption issue since the allocation
% of this resource was being verified very early in the application's flow, using the \texttt{tls\_get\_message\_header} function presented in Listing ~\ref{lst:vuln2} (\ref{lst:func3} and \ref{lst:func4}).
% These verifications of size allocation were substituted by the one presented in Listing~\ref{lst:vuln}(\ref{lst:func2}).
% %
% Although checking the maximum size of a message seems an elementary
% problem to repair, the example shows that there was a significant amount of
% changes performed in the code base which produced a negative impact in the
% maintainability. The developer tries to patch the
% code base by introducing a new branch point to the function \texttt{read\_state\_machine}~\ref{lst:func1} which adds complexity to the software
% and disrupts
% one of the guidelines proposed by the Software Improvement Group (SIG) for
% building maintainable software~\cite{Visser:2016:OREILLY} $-$
% the \emph{Write Simple Units of Code} guideline $-$ which resulted in a negative impact on the
% OpenSSL maintainability.}

%%%%%%% 
Software maintainability is designed as the degree to which an application is understood, 
repaired or enhanced. In this paper, our concern is to study whether, while improving software
security, developers are also negatively impacting the maintainability of their
software applications. This is important because software maintainability is 
approximately $75\%$ of the cost related to a project. To answer the following three 
research questions, we use two datasets of security patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064} 
to measure the impact of security patches on the maintainability of open-source software. 
%

\textit{\textbf{RQ1: What is the impact of security patches on the
maintainability of open-source software?}} Often security flaws require patching 
code to make software more secure.
However, \textbf{there is no evidence yet of how security patches impact the
maintainability of open-source software}. Our suspicion is that developers tend
to introduce technical debt in their software when patching the code base to
address a security flaw because they tend to choose the easiest path to solve
it. To address it, we compute the maintainability of $1335$ commits using the
\emph{Better Code Hub} tool. 
%

\textit{\textbf{RQ2: Which patterns of security patches are more likely to
affect open-source software maintainability?}}
There are security flaws that are more difficult to patch than others. For
instance, implementing secure authentication is not as easy as fixing a
cross-site scripting vulnerability since usually the last one can be fixed
without adding new lines of code/complexity to the code. A typical fix for 
the cross-site scripting vulnerability is presented in Listing~\ref{lst:fix}. 
The developer added the function \texttt{htmlentities} to escape the data given by the variable
\texttt{\$\_['file']}. \textbf{Knowing which patterns are more likely to increase
maintainability issues is one step forward to bring awareness to security
engineers of what patterns need more attention}. The taxonomy of security
patterns used to answer this question is the one provided by the Common Weakness Enumeration
(CWE). Maintainability is measured
separately for each pattern.
%
\setcounter{lstannotation}{0}
\begin{lstlisting}[style={PHPStyle}, caption={Fix provided by \texttt{nextcloud/server} developers to a \\Cross-Site Scripting vulnerability},label={lst:fix}]
   <p class='hint'>
    <?php
-   if(isset($_['file'])) echo $_['file']
+   if(isset($_['file'])) echo htmlentities($_['file'])
    ?>
   </p>
\end{lstlisting}
%
\textit{\textbf{RQ3: What is the impact of security patches versus regular changes on the
maintainability of open-source software?}}
%
We also compute the maintainability of random regular commits using the
\emph{Better Code Hub} tool --- baseline.  We use them to understand \textbf{how
maintainability evolves when security patches are performed versus when
they are not}.
%
\section{Methodology}\label{sec:methodology}
%
In this section, we discuss the methodology used to measure
the impact of security patches on the maintainbility of open-source
software. The methodology comprises the following steps, as illustrated in
Figure~\ref{fig:met}.
%
\begin{enumerate}
	\item Combine the datasets from related work that classifies
the activities of developers addressing security-oriented patches~\cite{Reis:2017:IJSSE, 10.1109/MSR.2019.00064}.
%
	\item
	Extract relevant data (e.g., owner and name of
	the repository, sha key of the vulnerable version, sha key of the fixed version) to the experiments and 
	code bases from the combined dataset containing $1335$ security
	refactorings collected from open-source software available on
	GitHub.
%
	\item A baseline of regular commits was randomly collected from the list of
	projects of the final dataset to evaluate the impact of regular commits on the
	maintainability of open-source software.
%
  \item Use the Software Improvement Group's\footnote{SIG's website: \url{https://www.sig.eu/} 
  (Accessed on \today{})} web-based source code analysis
  service \emph{Better Code Hub} (BCH)\footnote{BCH's website:
  \url{https://bettercodehub.com/} (Accessed on \today{})} to quantify maintainability
  for both security and regular commits. The code bases that resemble to the commits
  of our datasets are collected before the analysis performed by BCH.
\end{enumerate}
%
\begin{figure}[h]
 	\centering 	
	\includegraphics[width=0.5\textwidth]{figures/methodology.pdf}
 	\caption{Study Methodology}
	\label{fig:met}
\end{figure}
%
\subsection{Datasets}
%
We use a combined dataset of security vulnerabilities which is 
the outcome of mining and manually inspect a total of $265$ GitHub projects for both
datasets. Reis and Abreu 
($2017$) mined open-source
software aiming at the extraction of real examples --- created by real
developers --- to test and assess the performance of static analysis tools~\cite{Reis:2017:IJSSE} since
using hand-seeded test cases or mutations could lead to misleading assessments
of the capabilities of the tools~\cite{just2014mutants}. The study yielded a
dataset of $711$ test cases for $16$ different security vulnerability types. Some of the vulnerability types
are based on the OWASP Top $10$ of $2013$~\cite{oswap:2013} and OWASP Top $10$ of
$2017$~\cite{oswap:2017}. Each test case of the
dataset is a triplet: the commit before the patching, the commit responsible
for the patching, and the snippets of code that differ from one version to
another (typically, called \textit{diffs}) --- where one can easily review the
code used to fix the security flaw. 
%

Ponta et al. ($2019$) tracked the \url{pivotal.io} website for vulnerabilities 
from $2014$ to $2019$. For each new vulnerability, the authors manually searched 
for references to commits involved in the patch on the National Vulnerability 
Database (NVD) website. However, $70\%$
of the vulnerabilities did not have any references to commits. Thus, the authors
used their expertise to locate the commits in the repositories. This technique 
yielded a database of $624$ patches~\cite{10.1109/MSR.2019.00064}. The original 
dataset contains $1282$ commits. One patch can have multiple commits assigned.
To fit the dataset in our methodology, we search for the first and last commits
used to patch the vulnerability. We assumed the last commit of the patch as 
the fix (\emph{sha}) and the parent of the first commit as the vulnerable version 
(\emph{sha-p}).
%

In this study, we focus on computing the
maintainability of the commits before and after the security patching to
evaluate if its impact was positive, negative or none.
The $1335$ patches in the dataset were run against the BCH toolset to
calculate their maintainability reports. However, there are refactorings
discarded from our study due to limitations of BCH: in particular, lack of
language support and project size. The final dataset used in this paper comprises
$969$ security patches. $867$ of these patches are identified with Common Weakness
Enumeration IDs (CWE) where $62.1\%$ were scraped from the National Vulnerability 
Dataset (NVD) and the remaining $37.9\%$ were manually classified following the 
\emph{Research Concepts} CWE's list.

% In Table~\ref{tab:patterns}, the most prevailing categories are presented. To satisfy the Wilcoxon test size requirement, types of vulnerabilities
% with less than $20$ instances were merged in a bigger group named \textit{Miscellaneous}.

% \begin{table}[h]
% \scriptsize
% 	\caption{Guidelines to produce maintainable code}
% \begin{tabular}{L{1.1cm}L{2.1cm}L{4.3cm}}
%
% \toprule
% \textbf{Pattern} & \textbf{Name} & \textbf{Description}\\
% \midrule
%
% \textbf{CWE-89} &
%  Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases) &
%  When developers do not keep untrusted data
%      separate from SQL queries. If an attacker sends a command that
%      exploits the syntax of the SQL interpreter, then SQL injection attack is possible.\\\midrule
%
% \textbf{CWE-611} &
%  Improper Restriction of XML External Entity Reference (24 cases) &
% 	XML documents are exchanged through the web containing entities with URIs
%  	that resolve to local/external files. Thus, when XML parsers are not well configured,
%  	attackers have allowed to directly those files.\\\midrule
%
%  \textbf{CWE-20} &
%  Improper Input Validation (76 cases) &
% 	When developers
%  	forget to validate the inputs of an application, attackers may have control
%  	of the control/data flow of the program.\\\midrule
%
%   \textbf{CWE-200} &
% Information Exposure (39 cases)&
% Application's logs are one way of intentional/unintentional disclosure information
%  	to attackers. Many times attackers get access to logs when sould not be authorized to.\\\midrule
%
%    \textbf{CWE-352} &
% Cross-Site Request Forgery/CSRF (22 cases)&
% Poor session tokens generation
%      and management usually allow attackers to send forged HTTP requests
%      including authentication information from the victim to the vulnerable web
%      application.\\\midrule
%
%        \textbf{CWE-264} &
% Permissions, Privileges, and Access Controls (59 cases)&
% Incorrectly
%  	    implemented functionalities related to authentication and session
%  	    management, allowing an attacker to gain access to session tokens,
%  	    passwords, keys, and other sensitive data.\\\midrule
%
%  	           \textbf{CWE-399} &
% Resource Management Errors (22 cases)&
% Resource management issues are found more frequently
%  		in programming languages that do not manage memory automatically (e.g., C/C++
%  	    and Objective-C), i.e., where developers are responsible for
%  	    handling it instead. It is one of the main causes of Denial-of-Service attacks.\\\midrule
%
%  	     	           \textbf{CWE-79} &
% Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (63 cases)&
% Lack of proper validation or escaping
%  	    allows attackers to submit untrusted data to web browsers through malicious
%  	    scripts that can hijack the user sessions or redirect the user to malicious
%  	    sites.\\\midrule
%
%  	     	           \textbf{CWE-22} &
% Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases) &
% When developers forget to neutralize special elements within the path names to files/directories, attackers
%  	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
%  	or entire repositories.\\\midrule
%
%  	 	     	           \textbf{MISC} &
% Miscellaneous (666 cases) &
% This pattern comprises
%  		several other security refactorings that do not have a CWE assigned and patterns that
%  		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.\\
% \bottomrule
% \end{tabular}
% \label{tab:patterns}
% \end{table}

%
% \begin{itemize}
%     \item \textbf{CWE-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
%     separate from SQL queries. If an attacker sends a command that
%     exploits the syntax of the SQL interpreter, then SQL injection attack is possible.
% %
%     \item \textbf{CWE-611: Improper Restriction of XML External Entity Reference (24 cases).}
% 	XML documents are exchanged through the web containing entities with URIs
% 	that resolve to local/external files. Thus, when XML parsers are not well configured,
% 	attackers have allowed to directly those files.
% %
%     \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
% 	forget to validate the inputs of an application, attackers may have control
% 	of the control/data flow of the program.
% %
%     \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
% 	to attackers. Many times attackers get access to logs when sould not be authorized to.
% %
% 	\item \textbf{CWE-352: Cross-Site Request Forgery/CSRF (22 cases).} Poor session tokens generation
%     and management usually allow attackers to send forged HTTP requests
%     including authentication information from the victim to the vulnerable web
%     application.
% %
% 	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
% 	    implemented functionalities related to authentication and session
% 	    management, allowing an attacker to gain access to session tokens,
% 	    passwords, keys, and other sensitive data.
% %
% 	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
% 		in programming languages that do not manage memory automatically (e.g., C/C++
% 	    and Objective-C), i.e., where developers are responsible for
% 	    handling it instead. It is one of the main causes of DoS attacks.
% %
% 	\item \textbf{CWE-79: Improper Neutralization of Input During Web Page Generation/Cross-site Scripting (62 cases)} Lack of proper validation or escaping
% 	    allows attackers to submit untrusted data to web browsers through malicious
% 	    scripts that can hijack the user sessions or redirect the user to malicious
% 	    sites.
% %
% 	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers
% 	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
% 	or entire repositories.
% %
% 	\item \textbf{Miscellaneous (673 cases)} This pattern comprises
% 		several other security refactorings that do not have a CWE assigned and patterns that
% 		do not satisfy the Wilcoxon test size requirement of more than $20$ refactorings.
% \end{itemize}
%
\subsection{Security vs. Baseline Commits}
%
Previous studies attempted to measure the impact of regular refactorings on
open-source software maintainability~\cite{HEGEDUS2018313}. However, there is no
previous work focused on comparing the impact of security refactorings with
regular refactorings on maintainability.
We analyze the maintainability of regular commits (i.e., commits not related
to security fixes) and use them as a baseline.
The baseline dataset uses the security commits dataset as input.
Both datasets have the same size. For each
security commit, a commit with an approximate size is searched. Due to the complexity 
of some patches, it was not possible to find patches with the exact number of added and 
deleted lines. Our search uses the patch \emph{diff} to look for a new similar \emph{diff}. 
The algorithm integrates the following steps:

\begin{enumerate}
\item Calculates the \emph{diff} between the security patch commit (\emph{sha}) and its parent (\emph{sha-p}).
\item Selects a random commit from the project (\emph{sha-reg}).
\item Calculates the \emph{diff} between the random commit (\emph{sha-reg}) and its parent (\emph{sha-reg-p}).
\item Checks if the size from the regular change has the same/approximated size to 
the security patch.
\end{enumerate}

The pair of the random commit and its parent is accepted if the \emph{diff} 
size fits in the range size. This range widens every $10$ attempts to
search for a change with the same size. We originate the regular commits from the 
security commits to ensure that differences in maintainability are not a 
consequence of characteristics of different projects.

\subsection{Maintainability Analysis}

In this research, we follow a very similar methodology to the one presented 
in previous work on the maintainability of energy-oriented fixes~\cite{cruz2019energyoriented}. 
As said before, the web-based source code analysis service \emph{Better Code
Hub} (BCH) is used to collect the maintainability reports of the refactorings of
each project. Table~\ref{tab:guidelines} presents the $10$ guidelines $-$ proposed
by BCH's authors for delivering software that is not difficult to
maintain~\cite{Visser:2016:OREILLY} $-$ and maps each guideline to the metric 
calculated by BCH. Theses guidelines are calculated using the metrics 
presented in~\cite{criteria:2017} and also briefly explained on Table~\ref{tab:guidelines}. 
During each guideline evaluation, the tool determines the compliance towards 
one guideline by establishing limits for the percentage of code allowed to be 
in each of the $4$ risk severity levels
(\emph{low risk}, \emph{medium risk}, \emph{high risk}, and \emph{very high
risk}). If the project does not violate the thresholds, then it is compliant
with the guideline. These thresholds are calibrated by BCH using their own
data/experience --- using open-source and closed software systems. If a project is
compliant with a guideline, it means that it is at least $65\%$ better than the
software used by BCH to calculate the thresholds\footnote{Check the answer to
\emph{How can I adjust the threshold for passing/not passing a guideline?} at
\url{https://bettercodehub.com/docs/faq} (Accessed on \today{})}.

\begin{figure}[h]
 	\centering 	
	\includegraphics[width=0.44\textwidth]{figures/bch_report.png}
 	\caption{Maintainability report of OpenSSL's CVE-$2016$-$6304$ vulnerability
refactoring for the guideline \emph{Write Simple Units of Code} provided by
\emph{Better Code Hub}. This version of OpenSSL does not comply with the
guideline in the example since the bars do not reach the threshold points. This
example only complies with $\sfrac{1}{10}$ guidelines (\emph{Write Clean Code}).}
	\label{fig:bchrep}
\end{figure}

\begin{table}[h]
\scriptsize
	\caption{Guidelines to produce maintainable code}
\begin{tabular}{L{1.66cm}L{3.3cm}L{2.7cm}}

\toprule
\textbf{10 Guidelines} & \textbf{Description} & \textbf{Metric}\\
\midrule

\textbf{Write Short Units of Code} & Limit code units to $15$ LOCs because smaller
 units are easier to understand, reuse and test them & \textbf{Unit Size:} \% of 
 LOCs within each unit~\cite{criteria:2017} \\\midrule

\textbf{Write Simple Units of Code} & Limit branch points to $4$ per unit because
it makes units easier to test and modify & \textbf{McCabe Complexity:} \# of decision 
points~\cite{1702388,criteria:2017}\\\midrule

\textbf{Write Code Once} & Do not copy code because bugs tend to replicate at
multiple places (inefficient and error-prone) & \textbf{Duplication:} \% of redundant 
LOCs~\cite{criteria:2017}\\\midrule

\textbf{Keep Unit Interfaces Small} & Limit the number of parameters to at most
$4$ because it makes units easier to understand and reuse & \textbf{Unit Interfacing:} 
\# of parameters defined in a signature of a unit~\cite{criteria:2017} \\\midrule

\textbf{Separate Concerns in Modules} & Avoid large modules because changes in
loosely coupled databases are easier to oversee and execute & \textbf{Module Coupling:} \# of
incoming dependencies~\cite{criteria:2017} \\\midrule

\textbf{Couple Architecture Components Loosely} & Minimize the amount of code
within modules that are exposed to modules in other components & \textbf{Component Independence:} 
\% of code in modules classified as hidden~\cite{criteria:2017}\\\midrule

\textbf{Keep Architecture Components Balanced} & Balancing the number of
components ease locating code and allow for isolated maintenance & \textbf{Component Balance:} 
Gini coefficient to measure the inequality of distribution between components~\cite{criteria:2017} \\\midrule

\textbf{Keep your code base Small} & Reduce and avoid the system size because
small products are easier to manage and maintain & \textbf{Volume:} \# of LOCs converted 
to man-month/man-year~\cite{criteria:2017} \\\midrule

\textbf{Automate Tests} & Test your code base because it makes development
predictable and less risky & \textbf{Testability:} Ratings aggregation $-$ unit 
complexity, component independence and volume~\cite{Visser:2016:OREILLY}
 \\\midrule

\textbf{Write Clean Code} & Avoid producing software with code smells because
it is more likely to be maintainable in the future & \textbf{Code Smells:} 
\# of Occurrences~\cite{Visser:2016:OREILLY} (e.g., magic constants and long 
identifier names) \\
\bottomrule
\end{tabular}
\label{tab:guidelines}
\end{table}

Figure~\ref{fig:bchrep} shows an example of the report
provided by BCH for a project after finishing its evaluation. The example
refers to the OpenSSL CVE-$2016$-$6304$ vulnerability refactoring -
as described by Section~\ref{sec:motivation}. This
version of OpenSSL only complies with $1$ out of $10$ guidelines: \emph{Write
Clean Code}.

SIG defines \emph{Units} as the smallest groups of code that can be maintained
and executed independently~\cite{Visser:2016:OREILLY} (e.g., methods and
constructors in Java). One of the guidelines with which the project does not
comply is the one presented in the report (cf. Figure~\ref{fig:bchrep}):
\emph{Write Simple Units of Code}. BCH analyzes this guideline based on the
McCabe Complexity~\cite{1702388} to calculate the number of branch points of a
method. The bar at the bottom of the figure represents the top $30$ units that
violate the guideline, sorted by severity. The violation severities are
indicated using colors, and there is a legend to help to interpret them. The
green bar represents the number of compliant branch points per unit (\emph{at
most $5$}), i.e., the number of units are compliant with ISO
$25010$~\cite{iso:2011}. Yellow, orange and red bars represent units that do not
comply with medium (\emph{above $5$}), high (\emph{above $10$}) and very high
(\emph{above $25$}) severity levels. In the bar, there are marks that pinpoint
the compliance thresholds for each severity level. If the green mark is
somewhere in the green bar it means it is compliant with a low level of
severity.

\begin{figure}[h]
 	\centering 	
	\includegraphics[width=0.6\linewidth]{figures/commit.pdf}
 	\caption{Maintainability difference for security commits}
	\label{fig:commit}
\end{figure}

Aiming to analyze the impact of security patches, we use BCH to compute the
maintainability of two different versions of the project (cf. Figure~\ref{fig:commit}):

\begin{itemize}
	\item $v_{s-1}$, the version containing the security flaw, i.e., before the
	patch (\emph{sha-p});
	\item $v_{s}$, the version free of the security flaw, i.e., after the
	patch (\emph{sha});
\end{itemize}

Patches can be performed through one commit, several consecutive commits, or 
commit(s) interleaved with more programming activities, being the later
commonly called a floss-refactoring. This study considers mainly $1$-commit 
patches. Yet, a small percentage of data points had more than one commit 
involved in the patch. We manually inspected these cases and $23$ 
floss-refactorings were identified and disregarded since it would not
be fair to measure the maintainability of these cases where other
programming activities are involved.

The BCH tool does not compute the final score that our study needs to compare
maintainability amongst different project versions. For this, we follow previous 
work on measuring the impact of energy-oriented 
refactorings~\cite{cruz2019energyoriented}. Cruz et al. ($2019$) proposed an 
equation to capture the distance between the current state of the project and
the standard thresholds calculated by the BCH based on the insights provided 
in~\cite{Olivari:2018}. The equation provided in~\cite{cruz2019energyoriented} 
considers that the size of project changes does not affect the maintainability 
and that the distance to lower severity levels is less penalized than to the
thresholds in high severity levels.

% 	thresholds in high severity levels
% 	difference between two versions
% The equation considers the following:
% \begin{itemize}
% 	\item \textbf{Size of project changes does not affect the maintainability
% 	difference between two versions, $\Delta M (v_{s-1},v_{s}) = M(v_{s}) - M(v_{s-1})$.} We
% 	aim at evaluating security patterns occurring in different projects similarly.
% 	Thus, the derived metric uses the \textit{raw} number of lines of code rather
%   percentages for normalization purposes.
% 	\item \textbf{Distance to lower severity levels is less penalized than to the
% 	thresholds in high severity levels.} Severity level weights based on the
% 	severity level to count lines of code that violate maintainability guidelines.
% \end{itemize}

Given the violations for the BCH guidelines, the maintainability score is computed
$M(v)$ as follows:

\begin{equation}
    M(v) = \sum_{g \in G}^{} M_{g}(v)
\end{equation}

\noindent
where $G$ is the group of maintainability guidelines from BCH
(Table~\ref{tab:guidelines}) and $v$ is the version of the software under
evaluation. $M(v) < 0$ indicates that version $v$ is violating (some of) the
guidelines, while $M(v) > 0$ indicates that version $v$ is following
the BCH guidelines. The maintenance for the guideline $g$, $M_g$ for a given version
of a project is computed as the summation of the compliance with the 
maintainability guideline for the given severity level (medium, high, and very high).
The compliance for a severity level
is calculated based on previous work, which calculates the number of lines of code that comply and not comply
with the guideline at a given severity level~\cite{cruz2019energyoriented}. In our analysis, we compute 
the difference of maintainability between the security commit ($v_{s}$) and its parent 
commit ($v_{s-1}$), as illustrated in Figure~\ref{fig:commit}. Thus, we can determine 
which patches had a positive, negative or null impact on the project maintainability.

% The compliance $C$ for a given severity
% level $l$ is derived by:
%
% \begin{equation}\label{eq:3}
%     C(l) = LOC_{compliant}(l) - w(l) * LOC_{\neg compliant}(l)
% \end{equation}
%
% \noindent
% where $LOC_{compliant}(l)$ are the lines of code that comply with the guideline
% at the given severity level $l$, $LOC_{\neg compliant}(l)$ are the lines of code
% that do not comply with the guideline at the given severity level $l$ and $w(l)$
% is the weight factor to heighten the impact of non-compliant lines in comparison to
% compliant lines. Finally, the term $w(l)$ is calculated as follows:
%
% \begin{equation}
%     w(l) = \frac{1 - \theta(l)}{\theta(l)}
% \end{equation}
%
% \noindent
% where $\theta(l)$ is the threshold in percentage of the lines of code that are
% accepted to be non-compliant with the guideline for the severity level $l$. This
% is a standard value defined by BCH. In other words, the factor $w$ is used in
% Equation~\ref{eq:3} to highlight the lines of code that are not complying with
% the guideline. Then, we compute the difference of maintainability between the
% security commit ($v_{s}$) and its parent commit ($v_{s-1}$), as illustrated in
% Figure~\ref{fig:commit}.

\subsection{Statistical Validation}\label{sec:statsval}
%
To validate the maintainability differences in different groups of commits
(e.g., baseline and security commits), we use the Paired Wilcoxon signed-rank
test with the significance level $\alpha = 0.05$~\cite{10.2307/3001968}. In
other words, we test the null hypothesis that the maintainability difference
between pairs of versions $v_{s-1}$, $v_s$ (i.e., before and after a
security-commit) come from the same distribution. Nevertheless, this test 
has a limitation: it does not consider the groups of commits with a 
zero-difference maintainability. In $1959$, Pratt provided an improvement to 
the test to solve this issue making the test more robust. Thus, we use a 
version of the Wilcoxon test that incorporates the cases where 
maintainability is different from zero~\cite{10.2307/2282543}.
The Wilcoxon test requires a distribution size of at leat $20$ instances.
To understand the effect-size, as advocated by the Common-language effect 
sizes\cite{graw:1992}, we compute the mean difference, the median of the 
difference, and the percentage of cases that reduce maintainability.
%

\section{Results \& Discussion}\label{sec:results}

This study evaluates a final total of $969$ security and baseline commits 
from $260$ distinct open-source projects. The following section presents and 
discusses the results and answers to each research question.

\textit{\textbf{RQ1: What is the impact of security patches on the
maintainability of open-source software?}}

In \emph{RQ1}, we report and discuss the impact of patches on open-source
software maintainability under two different views: guidelines/metrics 
and formula proposed - the mean of all 
guidelines/metrics - $M(v)$.

\textbf{Guidelines/Metrics:} Each patch performs a set of changes
on the source code of the software. These changes may have a different
impact on the guidelines/metrics used to measure software maintainability. 
Figure~\ref{fig:guidelines} shows the impact of security patches 
on each guideline individually and the average impact on all guidelines 
together - $M(v)$. Under each guideline is stated the 
metric used to the calculations. Table~\ref{tab:guidelines} describes 
in more detail the metrics behind the guidelines presented. 
For each type of guideline, three types of impact of patches in 
the software maintainability are presented: 
\emph{negative}, maintainability decreases (\emph{red});
\emph{none}, maintainability remains the same (\emph{yellow}); and,
\emph{positive}, maintainability increases (\emph{green}). 
For each type of guideline, a boxplot 
and a swarmplot are presented to show the variability and dipersion of
the results alongside the number of absolute
and relative cases of each impact. Next to each
type of guideline, it is presented the mean ($\overline{x}$) and median (M)
of the maintainability difference and the p-value resulting from the
Paired Wilcoxon signed-rank test. $M(v)$ is not a guideline but rather the 
average impact of all guidelines.

Regarding the impact of security patches for each guideline, we observe
that overall patches have a more positive impact than negative in
all guidelines/metrics. However, patching vulnerabilities also has a very
significant negative impact on all guidelines - between $10\%$ and $40\%$. 
\emph{Write Short Units of 
Code} ($38.3\%$), \emph{Write Simple Units of Code} ($37.9\%$),  and \emph{Separate Concerns in Modules} ($33.8\%$) seem to be the 
most affected guidelines. 
This may imply that developers when patching 
vulnerabilities fail to design/implement solutions that continue to respect 
the limit bounds of branch points (/programming conditions) and function/module 
sizes (e.g., functions should not have more than $15$ LOCs) that are recommended
for coding best practices. Still, on 
respecting bound limits, developers also seem to not consider the limit 
of $4$ parameters per function for the \emph{Keep Unit Interfaces Small} 
guideline required by BCH, in $24.8\%$ of the cases. This guideline is usually
violated when the patch requires to input new information to a function/class 
and developers fail to use the \emph{Introduce Parameter Object} refactoring 
pattern. Results do not provide statistical significance to the \emph{Separate 
Concerns in Modules} guideline, i.e., results should be read carefully.

Software architecture is also affected while patching vulnerabilities.
Both \emph{Couple Architecture Component Loosely} and \emph{Keep
Architecture Components Balanced} guidelines suffer a negative impact of 
$26.9\%$ and $11.0\%$, respectively. Component independence and balance
are important to make easier to find the source code that developers
want to patch/improve and to understand how the high-level components
interact with others. However, results may imply that developers
fail to use techniques such as encapsulation to hide implementation
details and make the system more modular.

Software systems typically have $9\%$-$17\%$ of cloned code~\cite{5773403}. 
The \emph{Write Code Once} guideline results show that duplicated code 
increased in $171$ ($17.7\%$) of the patches. Previous work showed a 
correlation between code smells and code duplication~\cite{7476787} 
which may also be reflected in the \emph{Write Clean Code} guideline results. 
BCH reported new code smells for $145$ ($15.0\%$) patches which according 
to previous work may harm software security since some of these new code
smells may be new vulnerabilities~\cite{8819456}.
Thus, while developers 
tend to ignore refactoring techinques such as the \emph{Replace Method 
with Method Object} technique, they may also be introducing new software 
vulnerabilities and, consequently, harm the market value and economy of companies
~\cite{4267025}.

For projects with large codebases, the results retrieved for \emph{Keep Your 
Codebase Small} were not viable because they retrieved values above the limit 
set by BCH ($20$ Person-years). The \emph{Automated Tests} guideline was also 
not considered since the tool does not contemplate two
of the most important techniques to security testing: vulnerability scanning
and penetration testing. Instead, it only contemplates unit 
testing. Due to BCH limitations, we detected a few data points
that retrieved incorrect overall calculations. 
Those data points were disregarded from the study.

\textbf{Overall Score ($M(v)$):} 
Although overall patching vulnerabilities has a less negative impact
on software maintainability guidelines, this is not reflected in the average 
impact of all guidelines ($M(v)$) as we can see in Figure~\ref{fig:guidelines}.
The $M(v)$ plots show that $406$ ($41.9\%$) of the cases have negative impact
on software maintainability. While $188$ ($19.4\%$) have no impact and $375$ ($38.7$)
have a negative impact. The larger number of negative cases may be 
explained by guidelines with higher concentration of negative cases with higher amplitudes
such as WShortUC, WSimpleUC and SCM. In other words, it is 
important to note that the design debt of one guideline can lead
to severe impacts on the software quality~\cite{10.1145/1985362.1985366}.  
Thus, tools like Better Code Hub and static analysis tools (e.g., SonarQube, Codacy,
ESLint, Infer, and more) should be leveraged to 
assess the risk of patching vulnerabilities - alongside with pull
requests and code reviews. While static code analysis
may be daunting due to the number of rules and different effects
on maintainability, BCH only uses the top-$10$ of the guidelines
with the highest effect on maintainability. Static analysis 
can be used to complement the source code analysis, introduce
the capability of vulnerability detection, and, consequently,
reduce the risk of hindering maintainability.

 \begin{figure}[H]
  	\centering
  	\includegraphics[width=0.5\textwidth]{figures/main_per_guideline.pdf}
 	\vspace{-0.8 cm}
  	\caption{Impact of the security patches per guideline.}
 	\label{fig:guidelines}
 	\vspace{0.2 cm}
 	\scalebox{0.5}{
 	\begin{tabular}{r@{: }l r@{: }l}
 	\textbf{WCC} & Write Clean Code & \textbf{KACB} & Keep Architecture Components Balanced\\
 	\textbf{CACL} & Couple Architecture Components Loosely & \textbf{SCM} & Separate Concerns
	in Modules \\
 	\textbf{KUIS} & Keep Unit Interfaces Small & \textbf{WCO} & Write Code Once \\
 	\textbf{WSimpleUC} & Write Simple Units of Code & \textbf{WShortUC} & Write Short Units of
	Code
 	\end{tabular}
 	}
 	\vspace{-0.3 cm}
 \end{figure}

\textbf{Severity:} Some of the vulnerabilities are identified with \emph{Common 
Vulnerabilities and Exposure} (CVE) entries. We leveraged the \emph{National Vulnerability 
Database} (NVD) website to obtain their severity levels. In total, we 
retrieved severity scores for $536$ vulnerabilities: $112$ (\emph{HIGH}), 
$395$ (\emph{MEDIUM}) and $29$ (\emph{LOW}). Figure~\ref{fig:severity} 
presents the impact of security patches per severity level on the 
maintainability of open-source software. We observe that patches for 
\emph{HIGH} ($50.0\%$) and \emph{MEDIUM} ($43.8\%$) severity vulnerabilities 
hinder more the maintainability of software than \emph{LOW} ($27.6\%$) 
severity vulnerabilities. Again, patches have a considerable 
negative impact on software maintainability - between $20\%$ 
and $50\%$. Statistical significance was retrieved only for \emph{LOW}
severity vulnerabilitie, i.e., \emph{LOW} severity vulnerabilities
may have less negative impact on software maintainability thant the 
other severity levels. However, 
results should not be disregarded because they seem to confirm the 
assumption that higher severity vulnerabilities patches may have a 
more negative impact on maintainability and, consequently, that high/medium 
severity vulnerabilities may need more attention than low severity while 
patching vulnerabilities.

\begin{figure}[h]
 	\centering 	\includegraphics[width=0.4\textwidth]{figures/main_per_severity.pdf}
 	\caption{Maintainability difference by vulnerability severity.}
	\label{fig:severity}
\end{figure}

\textbf{Programming Languages:} The impact on software maintainability per programming 
language was also analyzed (Figure~\ref{fig:lang_main}). We restrict 
this analysis to programming languages with at least $20$ data points, as this 
is a requirement for our hypothesis tests (HT). Thus, we compare the results for 
\texttt{C/C++}, \texttt{Ruby}, \texttt{Java}, \texttt{Objective C/C++}, 
\texttt{Python} and \texttt{PHP}. \texttt{Groovy} samples are not presented
because the number of cases did not fulfill the HT requirement. 
\texttt{C/C++} ($46.5\%$), \texttt{Ruby} ($46.5\%$) and 
\texttt{PHP} ($38.6\%$) are the programming languages with worse 
impact on software maintainability. \texttt{Java} ($50.9\%$) and \texttt{Python} 
($46.5\%$) seem to be the 
languages less affected by patching, i.e., more than $45\%$ of patches had a positive 
impact on maintainability. But overall patches in all languages have a 
considerable negative impact on maintainability - between $35\%$ to $50\%$ - which 
confirms the need for better and more secure programming languages. 
Statistical significance was only retrieved for the \texttt{C/C++} 
(p-value = $2.24$x$10^{-05}$), \texttt{Ruby} (p-value = $0.041$) and \texttt{PHP} (p-value = $0.048$) languages.
Results should be read with care where statistical significance was not found.
Yet, data reports very interesting hints on the impact of programming languages 
on security patches.

We would expect that the negative impact of programming languages on
maintainability to be more severe, as arguably poor design of programming
languages and lack of using the best practices lead to more buggy and vulnerable
code~\cite{Ray:2017:LSP:3144574.3126905, 2019arXiv190110220B}. However,
Figure~\ref{fig:lang_main} shows that only \emph{C/C++} and \emph{Ruby} have 
a significant negative impact approximate to $50\%$ on software 
maintainability. We suspect that these values are
the result of projects contributions policies (e.g., coding standards). In our dataset, 
$9$/$10$ projects with more contributors have strict contribution 
policies regarding code standards. For
instance, \texttt{laravel/framework} requires the \texttt{PSR-2
Standard}\footnote{PSR-2 Standard available at
\url{https://www.php-fig.org/psr/psr-2/} (Accessed on \today{})} that is a coding style guide for \texttt{PHP}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.4\textwidth]{figures/main_per_language.pdf}
  \caption{Maintainability difference by programming language.}
  \label{fig:lang_main}  
\end{figure}

\textbf{Conclusion:} Results show that developers may fail to follow the coding
practice guidelines and hinder software maintainability after patching; and, 
that different levels of attention should be paid to each guideline. For instance,
\emph{Write Simple Units of Code} and \emph{Write Short Units of Code}
are the most affected ones. No statistical significance was retrieved
for \emph{Separate Concerns in Modules}. Different severity vulnerabilities
may need different levels of attention and better and more secure programming languages
are needed.
The resulting p-value of the Paired Wilcoxon signed-rank test for $M(v)$ in Figure~\ref{fig:guidelines}
is $0.044$. Since the p-value is 
below the significance level of $0.05$, we argue that \emph{security patches 
may have a negative impact on the maintainability of open-source software}; and, 
that tools such as Better Code Hub should be integrated in the CI/CD pipeline
to help developers evaluate the risk of patches hindering software maintainability
alongside with Pull Requests/Code Reviews.

% \Luis{I think this section is missing more discussion. What do you think should be changed in the community to cope with our findings? How can practitioners fix security vulnerabilities wihthout hindering maintainability? do we need more tools, better programming languages, more strict contribution policies? is there a trade-off between secure software and technical debt? show us how these findings can change SE}
% Following the \textit{American Statistical Association}
% recommendations~\cite{doi:10.1080/00031305.2016.1154108}, we consider the results for
% Ruby as statistically significant because its p-value is on the borderline.
%

\textit{\textbf{RQ2: Which weaknesses are more likely to
affect open-source software maintainability?}}

\begin{figure*}[htp]
  \centering
  \subfigure[Maintainability difference by first-level weaknesses from the 
  \textit{Research Concepts} list on Common Weakness Enumeration (CWE)]{
  \includegraphics[scale=0.4]{figures/main_per_cwe.pdf}}\quad\quad
  \subfigure[Maintainability difference by sub-weaknesses of the 
  \textit{Improper Neutralization} Weakness (CWE-707)]{
  \includegraphics[scale=0.4]{figures/main_per_cwe_spec_707.pdf}}\quad\quad
    \subfigure[Maintainability difference by sub-weaknesses of the 
	\textit{Improper Control of a Resource Through its Lifetime} Weakness (CWE-664)]{
	\includegraphics[scale=0.4]{figures/main_per_cwe_spec_664.pdf}}
	\label{fig:pat}
\end{figure*}

\emph{Weakness}, according to the Common Weakness Enumeration (CWE) glossary, 
is a type of mistake that, in proper conditions, could contribute to the 
introduction of vulnerabilities within that product. The CWE enumeration
system was used to classify the vulnerabilities included in our study.
Figure~\ref{fig:pat} shows three different charts: \emph{(a)} shows
the impact of all patches grouped using the first-level weaknesses from
the \emph{Research Concepts} list on CWE; and, the remaining charts
present the impact on maintainability of the sub-weaknesses from two of
the weaknesses presented in \emph{(a)}.

In Figure~\ref{fig:pat}-\emph{a}, we observe that patching vulnerabilities
such as Insufficient Control Flow Management 
(CWE-691), Protection Mechanism Failure (CWE-693), Improper Access Control 
(CWE-284), Improper Neutralization (CWE-707) and Improper Control of a Resource 
Through its Lifetime (CWE-664) may have a negative impact on software 
maintainability. The CWE-707 ($295$) and CWE-664 ($318$) 
weaknesses integrate a larger amount of cases compared to the remaining 
weaknesses. Thus, we present an analysis of their sub-weaknesses on 
Figure~\ref{fig:pat}-\emph{b} and Figure~\ref{fig:pat}-\emph{c}, respectively. 
Results demonstrate evidence that patching vulnerabilities hinders 
the maintainability of open-source software for $4$ different weaknesses: 
\emph{Improper Input Validation (CWE-20)}, \emph{Information Exposure 
(CWE-200)}, \emph{Missing Release of Memory after Effective 
Lifetime (CWE-401)} and \emph{Path Traversal (CWE-22)}. Results also show that 
software maintainability is less negatively impacted when patching 
\emph{Improper Restriction of XML External Entity Reference (CWE-611)}.

The impact of a patch depends on its complexity, i.e., if the patch
adds complexity to the code base it is probably affecting the software
maintainability. \emph{Cross-Site Scripting (CWE-79)} and 
\emph{Improper Restriction of Operations within the Bounds of 
a Memory Buffer (CWE-119)} patches endure more
cases with no impact on the open-source software maintainability.
Typically, CWE-79 vulnerabilities do not need extra lines 
to be fixed, as shown in Listing~\ref{lst:fix} - one simple 
\texttt{escape} function patches the issue. On the same type of fix,
CWE-199 vulnerabilities may also be fixed without adding new source code
(e.g., replacing the \texttt{strncpy} function by a more secure one 
\texttt{strlcpy} that checks if the buffer is null terminated). However,
some buffer overflows may be harder to fix and lead to more complex 
solutions such as the one performed for 
CVE-$2016$-$0799$\footnote{CVE-$2016$-$0799$ patch details available at 
\url{https://github.com/openssl/openssl/commit/9cb177301fdab492e4cfef376b28339afe3ef663} (Accessed on \today{})}.
As CWE-199 
vulnerabilities, \emph{Missing Release of Memory after Effective 
Lifetime (CWE-401)}, or also called \emph{Resource Management Errors (CWE-399)}, can also be the root cause of Denial-of-Service attacks and  
difficult to patch since it usually requires to add more complexity to the program 
- as seen in the example presented on Section~\ref{sec:motivation}. 
Patches of \emph{SQL Injection
(CWE-89)} weaknesses equally hinder and improve the software maintainability. 
But a very considerable percentage of patches hinder the open-source 
software maintainability ($41.1\%$).

\emph{Improper Input Validation (CWE-20)}, \emph{Information Exposure 
(CWE-200)}, \emph{Missing Release of Memory after Effective 
Lifetime (CWE-401)} and \emph{Path Traversal (CWE-22)} seem to be the most likely 
weaknesses to affect open-source software maintainability.
No statistical evidence was retrieved mostly due to the difficult
of finding a higher numbers of cases of open-source patches per weakness. However, 
results should not be disregarded because they offer an interesting view on how 
the maintainability is impacted per weakness which can provide insights to 
developers/researchers on the weaknesses that need more attention. Further
research should be perfomed to retrieve conclusions.

Real
efforts have been performed in creating and incorporating coding standards and best
practices in software production. However, security is still far from perfection
and fixing vulnerabilities might hinder software maintainability. Thus, it is
not only important to provide tool support to developers to help them apply these
patches without harming software maintainability, but also to create and redesign
programming languages to facilitate developers improving software maintainability.

\textit{\textbf{RQ3: What is the impact of security changes versus regular 
changes on the maintainability of open-source software?}}

The impact of security and baseline changes in the maintainability of
open-source software is presented in Figure~\ref{fig:secvsreg}.  
We have seen, previously, a deterioration on software maintainability 
when patching vulnerabilities: $41.9\%$ ($406$) suffered a 
negative impact, $38.7\%$ ($375$) remained the same and $19.4\%$ 
($188$) increased the software maintainability. For regular changes, 
we observe that the maintainability decreases in $27.0\%$ ($262$) 
and increases in $30.5\%$ ($295$) of the cases. But in contrast to 
security patches, the maintainability of regular changes remains the 
same in $42.5\%$ ($412$) of the cases and performing regular
changes has a more positive impact than negative on the software 
maintainability.

Results show that regular changes typically have less negative impact
on the software maintainability of open-source software.
%no impact on the software maintainability which contradicts previous research\Sof{missing citation}.
However, we manually inspected a considerable amount of cases from
the regular changes distribution where this happens and found that 
identifying regular commits as similar in size as the security-related 
commit is limiting the type of regular commits (e.g, input 
refactorings, variables or functions, type conversion, etc).  
We strongly believe that this phenomenon lead
to the significant number of cases where there is no impact
on the software maintainability.

For both types of changes, there is a negative impact
on software maintainability - especially, for security changes. This raises 
a new trade-off when developers need to patch vulnerabilities on their projects
because they might be hindering the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of security. We understand that developers
should be able to patch and produce secure code without hindering the maintainability of
their projects. But there are still some concerns that need to be addressed
and that this study tries to bring awareness for:

\begin{itemize}

	\item\textbf{Computer Curricula Needs to be Updated:}
	Software Developers lack the knowledge to create secure and
	maintainable code. Mainly because academic curricula is not yet prepared
	to properly educate them. Universities should do their part on making
	this possible. Online services such as BCH have also a great potential to help
	developers to be more aware of the maintainability issues introduced by their
	changes - alongside with pull requests and code reviews and integrated in 
	the CI/CD pipeline. Thus, developers can 
	put more work into improving their code maintainability and avoiding the risk
	of introducing new vulnerabilities after patching older ones.

	\item\textbf{Secure Programming Languages Wanted:} Programming Languages 
	should provide new design patterns to easily implement security patterns 
	without endangering maintainability.
	Ultimately, new programming languages, both secure and maintainable by design,
	may be designed to help developers be highly productive when writing and
	maintaining secure applications. One example is the need for designing new
	authorization mechanisms since these are one of the most complex security
	features to implement.

	\item\textbf{Poor Coding Practices Documentation:} 
	Software documentation must provided to developers with the
	best practices to implement secure and maintainable software. Unmaintainable
	code is difficult to test, analyze and re-use. Software companies must
	do an effort to provide these artifacts to their developers, in order,
	to avoid the introduction of new vulnerabilities while patching older ones.

\end{itemize}

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.4\textwidth]{figures/main_comparison.pdf}
 	\caption{Maintainability difference for security and baseline patches.}
	\label{fig:secvsreg}
\end{figure}

The resulting p-value of the Paired Wilcoxon signed-rank test is $0.331$, i.e., 
conclusions regarding the regular changes are not statistically significant. 
Yet, it is important to note that \emph{patching security vulnerabilities seem 
to have a more negative impact on the maintainability of open-source software 
than performing regular changes}.

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to the validity of this
study.
%
\\\textit{\textbf{Construct:}} The maintainability formula was inferred based 
on the BCH's reports. The high amount of different projects and backgrounds 
may require other maintainability standards. However, BCH does use a 
representative benchmark of closed and open-source software projects to compute 
the thresholds for each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012}.
Maintainability is computed as the mean of all guidelines. A patch can
have the same impact on the overall maintainability of both software versions 
- non-vulnerable and vulnerable - but not necessarily on the same 
guidelines.
%
\\\textit{\textbf{Internal:}} The security patches dataset provided by previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub
commits produced by project developers to classify the changes performed by them
to fix security flaws. This approach discards security patches that were
not explicit in commits messages. Pontas et. al\cite{10.1109/MSR.2019.00064} dataset
has patches that were manually mined in software repositories and, consequently,
wrong associations may have been performed between vulnerabilities and code snippets. 
The dataset also included floss-refactorings. Both cases were manually inspected
and deleted from the dataset. Thus, this study does not consider that a
developer may perform a patch quickly and later proceed to the refactor.
Baseline commits are retrieved randomly by selecting one commit with an 
approximate size and from the same project of a given security patch. 
This approach softens the differences
that may result from the characteristics of each project. However,
maintainability may still be affected by the developers' experience, coding
style and software contribution policies which is not evaluated in this study.
Furthermore, this evaluation considers that $969$ regular commits - any kind
of commit - are enough to
alleviate random irregularities in the maintainability differences of the
baseline. 
\\\textit{\textbf{External:}} We use a dataset that only comprises patches of open-source software.
However, the methodology requires to access data that is not publicly available.
Thus, our findings may not extend to non-open source software. Different programming 
languages may require different coding practices to address software safety. The 
dataset comprises more commits in Java, i.e., the dataset may not be representative 
of the population regarding programming languages. Our approach does not consider 
security patches in any other languages but English.

\section{Related Work}\label{sec:rw}

Many studies have investigated the relationship between refactorings and
software quality. Previous work focused on object-oriented metrics has evaluated the
impact of refactorings and exhibited proof that quantifying effectively the
impact of refactorings on maintainability may help to choose the appropriate
refactoring type~\cite{1167822}. In contrast to this work, Hegedus et
al.~\cite{HEGEDUS2018313} did not select particular metrics to assess the effect
of refactorings. Instead, statistical tests were used to find the metrics that
have the potential to change significantly after refactorings. The differences
of maintainability between groups of refactored elements and non-refactored were
analyzed. In addition, the authors concluded that the source code elements
subjected to refactorings had significantly lower maintainability than the
elements not affected. Studying the evolution of maintainability issues during
the development of Android apps, Malavolta et al. ($2018$)~\cite{8530041}
discovered that maintainability decreases over time. Palomba et al.
($2018$)~\cite{Palomba:2018:DIM:3231288.3231337} exhibits proof that code smells
should be carefully monitored by programmers since there is a high correlation
between maintainability aspects and proneness to changes and faults. In 2019, Cruz et 
al.~\cite{cruz2019energyoriented} proposed a formula to calculate maintainability 
based on the BCH's guidelines and measured the impact of energy-oriented fixes 
on the maintainability of software.
The present work uses the same model (BCH) and focuses solely on evaluating 
the impact of patching vulnerabilities on software maintainability.

Researchers investigated the relationship between design patterns and software
maintainability~\cite{10.1007/978-3-642-35267-6-18}. $300$ revisions of
\texttt{JHotDraw} were analyzed to conclude that every introduced pattern
instance improved software maintainability. However, other studies show that the
use of design patterns may introduce maintainability issues into
software~\cite{4493325}. Yskout et. al were not able to detect if the usage of 
design patterns has a positive impact, but concludes that developers prefer to 
work with the support of security patterns~\cite{8077802}. The present work 
studies how security patterns influence maintainability for open-source software.

There are studies that investigated the impact of programming languages on software
quality~\cite{Ray:2014:LSS:2635868.2635922,Ray:2017:LSP:3144574.3126905}. The first
study shows that some programming languages are more buggy-prone than others. However,
the authors of the second study were not able to reproduce it, not obtaining any
evidence about the impact of language design. Both have used the number of
defects as an indicator of software quality. In addition,
Berger et al. ($2019$)~\cite{2019arXiv190110220B} tried to reproduce~\cite{Ray:2014:LSS:2635868.2635922,
Ray:2017:LSP:3144574.3126905} and identified flaws that throw into distrust the 
previously demonstrated correlation between programming language and software 
defects. Our work studies how security refactorings affect software quality based 
on the code maintainability analysis and provides evidence that programming languages 
have an impact on maintainability.

\section{Conclusion and Future Work}\label{sec:conclusions}

This work presents an empirical study on the impact of $969$ security
patches on the maintainability of $260$ open-source projects. We leveraged
Better Code Hub reports to calculate maintainability based on a model proposed in previous work~\cite{Olivari:2018, cruz2019energyoriented}. Security patches 
significantly affect the software maintainability of open-source software, i.e., 
developers hinder maintainability while patching vulnerabilities. We found this evidence in
$41.9\%$ of the studied cases. When comparing with the impact of regular changes, security patches seem to have a higher negative impact than regular commits.
Regarding security patches, we also observe that some guidelines are more
affected than others: \emph{Write Simple Units of Code} ($37.9\%$) and
\emph{Write Short Units of Code} ($38.3\%$).

The current study also hints that more severe vulnerabilities may have
a more negative impact on the software maintainability and that patches
is some programming languages may needed more attention than others (e.g., \texttt{C/C++}, \texttt{Ruby} and \texttt{PHP}. Therefore, 
we draw attention for the need of better documentation explaining 
how to use best practices to produce secure and maintainable code; 
new and better programming languages; new
academic curricula in the fields of security and software quality; and, 
tools such as Better Code Hub and efficient static analysis to integrate
in the CI/CD pipelines to help 
developers predict the effect of their patches on software maintainability.

As future work, this study may be extended in several directions: 
investigate which guidelines affect most the maintainability per
weakness; check if vulnerability patches are followed by refactoring
commits and how much time does it take to do it; 
expand our methodology with other software quality properties; 
validate these findings with closed/private
software; and, expand this analysis to other quality standards.

%
% \section*{Acknowledgment}\label{sec:ack}
%
% \noindent
% We thank the Software Improvement Group (SIG) team for all the support and
% help in validating the methodology.
% %ack pedro adao when accepted.
% \pagebreak
\balance

{
  \bibliographystyle{unsrt}
  \bibliography{icsme20}
}

\end{document}