### 2021-04-30 response/response_2/response.tex
adding final review draft

\documentclass[11pt,fleqn]{article}

\usepackage{vmargin}
\setpapersize{A4}
\setmarginsrb{2.5cm}{2.5cm}{2.5cm}{2.5cm}%
{\baselineskip}{2\baselineskip}{baselineskip}{2\baselineskip}
\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

\usepackage{caption}
\usepackage{subcaption}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{latexsym}
\usepackage{url}
\usepackage{tgheros}
\renewcommand*\familydefault{\sfdefault}

\newcommand{\eline}{\vspace*{.75\baselineskip}}
\newcommand{\Referee}[1]{\eline \noindent {\bf Reviewer comment #1:} \\}
\newcommand{\Us}{\eline \noindent {\bf Response:}\\}
\newcommand{\TBD}{{\bf To Be Done}}
\newcommand{\newreviewer}[1]{\section*{Reviewer #1}\vspace*{-1.05\baselineskip}}
\newcommand{\editor}[1]{\section*{Editor #1}\vspace*{-1.05\baselineskip}}

%\usepackage{tikz}
\usepackage[skins,breakable]{tcolorbox}
\tcbset{textmarker/.style={
%
skin=enhancedmiddle jigsaw,breakable,parbox=false,before skip=-1mm, after skip=0mm,
boxrule=0mm,leftrule=2mm,rightrule=2mm,boxsep=0mm,arc=0mm,outer arc=0mm,
left=2mm,right=2mm,top=2mm,bottom=2mm,toptitle=1mm,bottomtitle=1mm,oversize}}

\newtcolorbox{rcomment}{textmarker,colback=gray!15!white,colframe=gray!80!white}

\newenvironment{revcomment}[1][]
{\Referee{#1}\begin{rcomment}}
{\end{rcomment}}

\newenvironment{reveditor}
{\begin{rcomment}}
{\end{rcomment}}

\usepackage{todo}
\usepackage{hyperref}

%customized
\usepackage{xfrac}

% more lenient line breaking (avoids text protruding into margins)
\sloppy

\title{\vspace*{-2cm}{\bf Authors' Response to the Review of
 EMSE-D-20-00300:\\
 ``Fixing Vulnerabilities Potentially Hinders Maintainability''}}

\author{Sofia Reis, Rui Abreu, Luis Cruz}
\date{}

\begin{document}

\maketitle

\editor{}

\begin{reveditor}
    Many thanks for submitting your revision to Empirical Software Engineering journal. 
    All reviewers and I thank you for the effort you have put in preparing the detailed 
    revision which successfully addresses all points raised earlier. Accordingly, I am happy 
    to recommend the acceptance of your manuscript to the EiCs. Please note that reviewers 
    highlight some editing issues, so please clarify Reviewer \#2's points and conduct an 
    overall proof read in preparing your final proofs.

\end{reveditor}

\Us We thank the editor and reviewers for their valuable feedback and acceptance. In the 
following, we answer the questions raised by the reviewer 2.

\newreviewer{2}

\begin{revcomment}[2.1]
    What is meant by "20 person years"?  How would one obtain this measure?
\end{revcomment}

\Us A person-year is the equivalent to 12 person-months. A person-month is a standard measure of 
source code volume where the total volume in a codebase is the volume in lines of code
converted to person-months. BCH uses person-years instead of volume 
because it facilitates the comparisons of source code volume between technologies
(different programming languages have different levels of of verbosity). 
We do not consider this metric on our study. However, according to the "Building
Maintainable Software" book, 20 person-years of rebuild value 
is the equivalent to at most 1750000 lines of code in Java.

\begin{revcomment}[2.2]

    The bullet for automated tests.  Can you think of another work 
    instead of "contemplate"?  I'm not sure of what you are trying to say.

\end{revcomment}

\Us Thank you for pointing this out. We changed the work for ``consider''.
This metric only considers unit testing, which is usually not applied for 
security testing but rather to functionality testing.

\end{document}